---
title: "Day5_Activity_Akhavi"
author: "Arash Akhavi"
date: "10/1/2021"
output: 
  prettydoc::html_pretty:
  theme: cayman
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(warning = FALSE, message = FALSE) 
```

```{r libraries}
library(tidyverse)
library(reprex)
```

```{r}
#Create a function called times_seven and multiply a vector of 7,8,9 by the function
#Create custom error messages so user knows what value to input
times_seven <- function(x) {
   if (!is.numeric(x)) {
     stop("Please enter a number!")
   } 
  {
   if (x ==7)
     cat("Sevens <3")
   }
	return(x * 7)
}

x <- c(7,8,9)
times_seven(x)
```
1. Write a function called divide_and_round(), which takes a number or vector of numbers, divides them all by their smallest value, and rounds the results to the nearest whole number.
```{r}
divide_and_round <- function(nums) {
  my_min <- min(nums, na.rm = TRUE)
  divided <- nums/my_min
  divided1 <- round(divided)
  return(divided1)
}

test <- c(1:5, NA)
divide_and_round(test)
```
2. Write a function called no_nines_or_twelves(), which takes a number or vector of numbers and returns TRUE if the number is NOT divisible by 9 or 12, FALSE otherwise.
```{r}
no_nines_or_twelves <- function(nums) {
   check <- !(nums %% 9 == 0 | nums %% 12 == 0)
   return(check)
}

nums <- c(7, 9, 12, 108)
test <- c(1:10*5, NA)
no_nines_or_twelves(test)
```
Fibonacci Example
```{r}
fibonacci <- function(start, len) {
  fib <- vector()
  fib[1:2] = 1
  for (i in 2:len) {
    fib[i + 1] <- fib[i] + fib[i - 1]
  }
  return(fib)
}

fibonacci(1,7)
```
3. Write a function called every_other() which takes a vector and returns every other value in the vector. Accomplish this with a for loop.
Include an optional argument called “start” which lets you choose where to start skipping; that is, if start = 1, we return the 1st value, 3rd, 5th, etc. and if start = 2, we return the 2nd, 4th, 6th, etc.

```{r}
every_other <- function(vec, start) {
  x <- vector()
  for(i in start:length(vec)) {
    if(start %% 2 == 0) {
      if(i %% 2 == 0) {
        x <- append(x, vec[i])
      }
    } else if(start %% 2 != 0) {
      if(i %% 2 != 0){
        x <- append(x, vec[i])
      }
    }
  }
  x
}

test <- c(1:10)
every_other(test, start = 2) 
```


4. Write a function called shorten() which takes a vector and keeps dropping the first value, until the sum of the vector is less than 350 Return the remaining values.

```{r}
shorten <- function(my_vec) {
while (sum(my_vec) >= 350) {
  my_vec <- my_vec[-1]
}
  my_vec
}
my_vec <- 1:50
shorten(my_vec)
```
5.Once you have written your four functions correctly, copy-pase and run the following code:
```{r}
my_vec <- c(39, 1.87, 48, 11, 8, 45, 21, 5, 12, 33, 9, 11, 108, 4, 18, 5, 16, 17, 8, 48, 27, 24, 4, 22, 12, 27, 23, 46, 42, 35, 15, 34, 36, 26, 18, 10, 18.21, 72.04, 36.9, 41.81, 29, 89.75, 34.03, 20.18, 48.74, 15.76, 31.86, 83.6, 43.55, 39.99, 23.55, 8.54, 24.71, 22.02, 9.71, 62.14, 35.46, 16.61, 15.66, 21.29, 30.52, 201.07, 45.81, 7.85, 30.13, 34.14, 22.62, 10.2, 6.02, 30.12, 10.64, 31.72, 24.57, 14.43, 43.37, 89.93, 44.72, 51.32, 13.62, 45.56, 22.96, 7.05, 29.99, 41.38, 26.59, 23.04, 19.82, 50.73, 39.56, 43.79, 30.22, 85.85, 5.78, 78.85, 29.52, 66.27, 44.06, 27.28, 24.43, 64.32, 3.35, 67.45, 46.72, 48.44, 48.65, 33.3, 40.28, 19.04)

my_vec <- every_other(my_vec, start = 2)
my_vec <- divide_and_round(my_vec)
my_vec <- every_other(my_vec, start = 1)
my_vec <- shorten(my_vec)
my_vec <- my_vec[no_nines_or_twelves(my_vec)]

my_vec <- sort(my_vec)

my_vec
```

