---
title: "Lab2_Akhavi"
author: "Arash Akhavi"
date: "10/1/2021"
output: 
  prettydoc::html_pretty:
  theme: cayman
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(warning = FALSE, message = FALSE) 
```

```{r libraries}
library(prettydoc)
library(tidyverse)
library(reprex)
```
1.
```{r}
#Import Data
hiphop <- read_csv("Data/hiphop.csv")
hiphop
```
This data set provides information about the familiarity of different words used in hip hop songs. A sample size of participants were surveyed regarding their familiarity with various words. Their demographic information such as age, gender, ethnicity, sex were recorded. Additionally, their familiarity with certain artists were recorded. 

2. Clean the dataset in whichever ways you see fit. This might mean adjusting variable type, for example from “character” to “factor”, or dealing with missing data.

```{r}
hiphop1 <- hiphop %>% 
  mutate(
    word = as.factor(word),
    subj = as.factor(subj),
    sex = as.factor(sex),
    ethnic = as.factor(ethnic)
  )

hiphop1
```
# 3. How many unique AAVE words were studied in this dataset?
```{r}
length(unique(hiphop$word))
```
4. Make a new variable that recategorizes ethnic into only two groups, “white” and “non-white”, to simplify your data
```{r}
hiphop_sorted <- hiphop1 %>% 
  mutate(
    white = ifelse(ethnic == "white", 1,0)
    )
hiphop_sorted
```

5. What are the demographics of the people in this study? Investigate the variables sex, age, and ethnic and summarize your findings in 1-3 complete sentences.
```{r}
hiphop_sorted %>% 
  select(age, sex, ethnic) %>% 
  summarize_all(c("n_distinct", "mean"))
```
In this data set there are nine total distinct ethnicities, 21 distinct ages, and two distinct sexes. Of the range of 21 ages the average age in the set of data is 20 years old.

6. Make at least two plots to display the demographic information of the subjects in this study.  but make sure they are appropriate to the data types and have informative titles and axis labels
```{r}
hiphop_sorted %>% 
  filter(word=="5-0") %>% 
  ggplot(aes(x = age, fill = sex)) +
  geom_bar() + 
  xlab("Age") + 
  ylab("Count") + 
  ggtitle("Ages of Males and Females")
```
```{r}
hiphop_sorted %>% 
  filter(word=="5-0") %>% 
  ggplot(aes(x = ethnic, fill = sex)) +
  geom_bar() + 
  xlab("Ethnic") + 
  ylab("Count") + 
  ggtitle("Ethnicities of Males and Females")
```
#  Familiar Words
1. For each demographic group listed below, determine which word(s) in this study was the most and least familiar on average.
a. People below the age of 20
```{r}
hiphop2 <- hiphop_sorted %>% 
  select(word, white, sex, age, familiarity) %>% 
  filter(age < "20") %>% 
  group_by(word) %>% 
  summarize(mean_familiarity = mean(familiarity))

hiphop2 %>% 
  arrange(mean_familiarity, desc = TRUE) %>% 
  top_n(1, mean_familiarity)

hiphop2 %>% 
  arrange(mean_familiarity, desc = TRUE) %>% 
  top_n(-1, mean_familiarity)
```
b. Non-white women
```{r}
hiphop3 <- hiphop_sorted %>% 
  select(word, white, sex, age, familiarity) %>% 
  filter(white == 0, sex == "Female") %>% 
  group_by(word) %>% 
  summarize(mean_familiarity = mean(familiarity))

hiphop3 %>% 
  arrange(mean_familiarity, desc = TRUE) %>% 
  top_n(1, mean_familiarity)

hiphop3 %>% 
  arrange(mean_familiarity, desc = TRUE) %>% 
  top_n(-1, mean_familiarity)
```
c. White men above the age of 30
```{r}
hiphop4 <- hiphop_sorted %>% 
  select(word, white, sex, age, familiarity) %>% 
  filter(white == 1, sex == "Male", age >= "30") %>% 
  group_by(word) %>% 
  summarize(mean_familiarity = mean(familiarity))

hiphop4 %>% 
  arrange(mean_familiarity, desc = TRUE) %>% 
  top_n(1, mean_familiarity)

hiphop4 %>% 
  arrange(mean_familiarity, desc = TRUE) %>% 
  top_n(-1, mean_familiarity)
```
2. For each demographic comparison below, determine which music genre most differentiates the groups. That is, which genre had much higher average (mean or median) score in one group than the other.

a. Male versus Female
```{r}
hiphop_sorted %>% 
    select(sex, intl, vocal, classical, folk, rock, country, pop, alternative, hiphop, unclassifiable) %>% 
  filter(sex == "Female") %>% 
  summarize_each(mean)- 
  hiphop_sorted %>% 
  select(sex, intl, vocal, classical, folk, rock, country, pop, alternative, hiphop, unclassifiable) %>% 
  filter(sex == "Male") %>% 
  summarize_each(mean)
```
In the male vs female group the genre rock had the biggest difference of -1.1272. 

b. White versus Non-White
```{r}
hiphop_sorted %>% 
    select(white, intl, vocal, classical, folk, rock, country, pop, alternative, hiphop, unclassifiable) %>% 
  filter(white == "1") %>% 
  summarize_each(mean)- 
  hiphop_sorted %>% 
  select(white, intl, vocal, classical, folk, rock, country, pop, alternative, hiphop, unclassifiable) %>% 
  filter(white == "0") %>% 
  summarize_each(mean)
```
In the white vs non-white group the genre international had the largest difference of -1.592593.

c. Age below 21 versus age 21+
```{r}
hiphop_sorted %>% 
    select(age, intl, vocal, classical, folk, rock, country, pop, alternative, hiphop, unclassifiable) %>% 
  filter(age >= "21") %>% 
  summarize_each(mean)- 
  hiphop_sorted %>% 
  select(age, intl, vocal, classical, folk, rock, country, pop, alternative, hiphop, unclassifiable) %>% 
  filter(age < "21") %>% 
  summarize_each(mean)
```
In the age below 21 and the age above 21 group the genre of music with the largest difference was alternative with a difference of 0.380597.

#  Use the data
1. Aubrey hopes that his songs will be percieved as authentically hiphop. He hopes his lyrics will be recognizeable to those who describe themselves as hiphop fans, but less recognizeable to those who do not consider themselves fans. Suggest some words or phrases that Aubrey should try to use, and some words he should avoid
```{r}
#Hiphop fans
hiphop_sorted %>% 
  select(age, word, hiphop, familiarity) %>% 
  filter(hiphop > 4) %>% 
  group_by(word) %>% 
  summarize(
    fans <- mean(familiarity)
  ) %>% 
    arrange(desc("fans"))

#Not fans of hiphop
hiphop_sorted %>% 
  select(age, word, hiphop, familiarity) %>% 
  filter(hiphop <= 2) %>% 
  group_by(word) %>% 
  summarize(
    fans <- mean(familiarity)
  ) %>% 
    arrange(desc("fans"))
```
Aubrey should use words and phrases such as "off the hook", "make it rain", "hella", "feel me". He should avoid phrases such as "wile out", "road dog", "plex", and "mail". 

2. Although Aubrey wants to be authentic, he also hopes to sell records, of course. Two titles have been suggested for his first album: “Hotline Boo” or “Hella Bling”. Based on the dataset, which will appeal more to the higher population areas? Make at least one plot to support your answer
```{r}
titles <- hiphop_sorted %>% 
  select(age, word, hiphop, familiarity, city, county) %>% 
  filter(city > 150000,
         word == c("hotline", "boo", "bling", "hella")) %>% 
  group_by(word) %>% 
  summarize(
    fans = mean(familiarity)
  ) 

titles %>% 
ggplot(aes(x = word, y = fans, fill = "blue2")) + 
  geom_bar(stat = "summary", color = "black") + 
  xlab("Word") + 
  ylab("Average Familiarity") + 
  ggtitle("Average Familiarity of Hiphop Words in Populated Cities")
```
In higher population areas of cities with greater than 150,000 residents familiarity with "hella" is high thereby "Hella Bling" would be a wiser choice as it is more popular. 

3. Aubrey’s true life dream is to collaborate with his fellow Canadian musician Justin Bieber. Luckily, he knows that Bieber himself was one of the subjects in this study! You know that Bieber is a white male, aged 17-23 at the time of the study, from a relatively small town (10,000-60,000 people) in Ontario.

Determine which subject is secretly Bieber, and justify your answer.
```{r}
justin <- hiphop_sorted %>% 
  select(subj, age, word, white, sex, hiphop, familiarity, city, county, bieber) %>% 
  filter(city >= 10000 & city <= 60000,
         age >= 17 & age <= 23,
         white == 1,
         sex == "Male",
         bieber >= 5)
justin
```
Justin Bieber is undercover as p17. Based on all of the filters applied, knowledge of Bieber songs, age between 17-23, city population between 10,000-60,000, and white male this is likely Bieber. 

