---
title: "Day8_WebScraping_Activity_Akhavi"
author: "Arash Akhavi"
date: "10/14/2021"
output: 
  prettydoc::html_pretty:
  theme: architect
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(warning = FALSE, message = FALSE) 
```

```{r}
library(tidyverse)
library(stringr)
library(rvest)
library(dplyr)
```

Web scrapping Example 1
```{r}
#Web scrape San Luis Obispo weather data from desired URL 
#Run html_table and html_text to see which works best
weatherurl <- "https://forecast.weather.gov/MapClick.php?lat=35.2603&lon=-120.6946#.YWhccRrMI2w"
weatherlink <- read_html(weatherurl)
forecasthtml <- html_nodes(weatherlink, css = ".forecast-text , #detailed-forecast-body b")
forecasttext <- html_text(forecasthtml)
paste(forecasttext, collapse = " ")

#Create a data frame with combined data
forecastdf <- data.frame(
  time_of_day = forecasttext[seq(from = 1, to = 26, by = 2)],
  forecast = forecasttext[seq(from = 2, to =26, by =2)]
)
```

Activity
```{r}
#Scrape the data from the crypto currency webpage
cryptourl <- "https://goldprice.org/cryptocurrency-price"
cryptolink <- read_html(cryptourl)
cryptohtml <- html_nodes(cryptolink, css = ".views-field-field-crypto-price-change-pc-24h , .views-field-field-crypto-price , .views-field-field-market-cap , .views-field-field-crypto-proper-name a")
cryptotext <- html_text(cryptohtml)


#Clean data
clean <- str_replace_all(cryptotext, "\\$", "") 
clean1 <- str_replace_all(clean, "\n", "")
clean2 <- str_replace_all(clean1, "%", "")
clean3 <- str_replace_all(clean2, ",", "")

#Create a data frame with the scrapped data
cryptodf <- data.frame(
  crypto_currency = clean3[seq(from = 1, to = 260, by = 4)],
  market_cap = clean3[seq(from = 2, to = 260, by = 4)],
  price = clean3[seq(from = 3, to = 260, by = 4)],
  daily_change_percent = clean3[seq(from = 4, to = 260, by = 4)]
)


#Mutate daily change column from percents to decimals
cryptodf1 <- cryptodf %>% 
  mutate(
    daily_change_decimals = (as.numeric(daily_change_percent)) / 100,
    market_cap_num = (as.numeric(market_cap)),
    price_num = (as.numeric(price))
  )
cryptodf1

#Identify top crypto currency with regard to Market Cap
top_market_cap <- cryptodf1 %>%                                     
  arrange(desc(market_cap_num)) 
top_market_cap  

#Identify top crypto currency with regard to Price 
top_price <- cryptodf1 %>%                                     
  arrange(desc(price_num))
top_price  

#Identify top crypto currency with regard to Change 
top_daily_change <- cryptodf1 %>%                                     
  arrange(desc(daily_change_decimals))
top_daily_change  
```
The top two crypto currencies with the highest market cap values are Bitcoin and Etherium with market caps of respectively $1,082,695,399,376 and $447,511,744,791.

The top two crypto currencies with the highest prices are Bitcoin and  year.finance with prices of respectively $57,450 and $35,240.

The top two crypto currencies with the highest daily change values are Aave [old] and Hedera with daily change values of respectively 0.5113 and 0.0741.

 
